Propaganda Films and National Identity: Animated Films in USSR and China

### 2025-03-10 paper/paper.qmd
finish 1930s soviet history in backgrounds section

The Cold War is a geopolitical and ideological war that defines global politics between the United States and the Soviet Union emerged after World War II from 1947 to 1991. The conflict between the two countries extends beyond the military advancement, reaching political, economical, and cultural areas [@riabovImagesUrbanSpace2021, pp. 123]. Animated films, as one of the cultural products, often include the use of propaganda as a tool to shape public opinion and assert ideological dominances [@pontieriSovietAnimationThaw2012, pp. 6]. The USSR and its allies promoted communism as a superior alternative to capitalism, which is dominant in the United States. Both the USSR and the newly established People’s Republic of China needs cultural products to prove the rightness of their ideological system. During the 1950s and 1960s, both the USSR and China utilize cinema as a medium for propaganda to gain public support for their own views on socialism. In this essay, we compare six propaganda animated films (three from the USSR and three from PRC) from the two countries in the 1950s and 1960s. We aim to identify the differences between these films and the social-political reasons behind them. First, we introduce the historical background in the post WWII era in both countries. Second, we introduce the six films and the directors. This includes the director's choice of style across different films, depictions of characteristics including movement, background, and sound, and the overall theme, objective, and target audiences. Finally, we connect the characteristics observed in these films and make connections with the general trends in the local animation industry, as well as historical and socio-cultural context. We conclude that ...

@@

>>>>

The Cold War is a geopolitical and ideological war that defines global politics between the United States and the Soviet Union emerged after World War II from 1947 to 1991. The conflict between the two countries extends beyond the military advancement, reaching political, economical, and cultural areas [@riabovImagesUrbanSpace2021, pp. 123]. Animated films, as one of the cultural products, often include the use of propaganda as a tool to shape public opinion and assert ideological dominances [@pontieriSovietAnimationThaw2012b, pp. 6]. The USSR and its allies promoted communism as a superior alternative to capitalism, which is dominant in the United States. Both the USSR and the newly established People’s Republic of China needs cultural products to prove the rightness of their ideological system. During the 1950s and 1960s, both the USSR and China utilize cinema as a medium for propaganda to gain public support for their own views on socialism. In this essay, we compare six propaganda animated films (three from the USSR and three from PRC) from the two countries in the 1950s and 1960s. We aim to identify the differences between these films and the social-political reasons behind them. First, we introduce the historical background in the post WWII era in both countries. Second, we introduce the six films and the directors. This includes the director's choice of style across different films, depictions of characteristics including movement, background, and sound, and the overall theme, objective, and target audiences. Finally, we connect the characteristics observed in these films and make connections with the general trends in the local animation industry, as well as historical and socio-cultural context. We conclude that ...

@@

The beginning of propaganda films in the USSR begins in the 1920s and 1930s, according to Eleanor Cowen and Laura Pontieri. The Stalin communist regime, they are particularly interested in the indoctrination of its youth. In 1936, they form a newly collectivized studio called Soyuzmultfilm in Moscow. The studio becomes the state-controlled premier animation studio in the Soviet Union [@cowenAnimationIronCurtain2020, chap. 4]. The early years of Soyuzmultfilm included a great deal of pro-Soviet and anti-capitalist propaganda, and they use Disney’s full, cel animation method as the blueprint. The aesthetic goal of complete animation is to create animations with great attention to detail and subtle fluidity of motion. This aligns with Josef Stalin and Maksim Gorki’s promotion of socialist realism ideology which focuses on optimistic and realistic representations of the country’s glorious future [@cowenAnimationIronCurtain2020, chap. 4]. The stories focus on banishing old bourgeoisie, and attacking the rich, racist and duplicitous capitalists [@pontieriSovietAnimationThaw2012b, p. 21, 36].
